---
title: "tutorial_2"
author: "Dani Chu"
date: "15/10/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
library(snakecase)
```

Some code from last week. Make sure you have a folder called data that has the data called `train.csv`.

```{r}
data <- 
  read_csv("data/train.csv", col_types = list(WindSpeed = col_character()))

data <-
  data %>%
  rename_all(to_snake_case)
```


# select function

the `select` function allows us to "select" columns. 


So here is our data 

```{r}
data
```

Now if we only want the columns witht the x and y location of the player and game play and player identifiers.

```{r}
data %>%
  select(x, y, play_id, game_id, nfl_id)
```

Now with that example select all the columns that are player specific.

```{r}
# add in your own code
data
```


# filter function

Let's say we want to find all the plays where Todd Gurley is on the field. We want to "filter" our data set for when we see Todd Gurley. There are 5 main conditions, equality `==`, less than `<`,  greater than `>`, less than or equal to `<=`, greater than or equal to `>=`

```{r}
data %>%
  filter(display_name == "Todd Gurley")
```

Now we have way too many columns so let's use what we learned before to only get the `play_id` column. Save that into a variable called `todd_play_ids`

```{r}
todd_play_ids <-
  # add in your own code
  data 
```

Now we want to have all the players on the field on plays where Todd Gurley was on the field. We're going to use a new condition called `%in%`. This says if the value in our column is "in" the "column" on the right then keep it.

```{r}
data %>%
  filter(play_id %in% todd_play_ids$play_id)
```

Finally, try this yourself. We want a dataset that has all the players who are in the end zone at the time of handoff. One endzone is between 0 and 10 the other one is between 110 and 120.

```{r}
# add in your own code
data

```

# mutate function

The mutate funciton allows us to create new columns.

# group by / summarise functions


